Unbalanced Mallows Models for Optimizing Expensive Black-Box Permutation Problems

### 2021-02-04 paper/main.tex
sec 5.2

\usepackage{booktabs} % For formal tables
\usepackage{enumitem}
\usepackage{afterpage,tabularx}
%\usepackage[colorlinks=true, allcolors=blue]{hyperref}

@@

\subsection{Rankings and orderings: experiments}\label{sec:exper_ro}
\ekhine{}{este primer parago es bastante importante para dejarlo aqui. }
In Section~\ref{sec:preliminaries} we highlighted the great difference of rankings and orderings, usually ommitted in the literature. In this lines, we show the consequences of misinterpreting permutations. We remind the reader that statistical models (Borda or the \ken distance) are defined for rankings while standard literature in optimization considers orderings. This is the reason while in the optimization field we will usually find the Swap distance instead of the \ken distance. However, both are usually claimed to be the same, which is false: they are the same in particular cases, for example when  one of the permutations is the identity or an inversion.

@@

  %% MANUEL: Ya dimos la URL arriba, así que no la repetiría aquí
  % \footnote{\url{http://grafo.etsii.urjc.es/optsicom/lolib/}}
%The type of instances differs greatly from the synthetic LOP instances, and,
%thus, so does the landscape. However, UMM shows a very similar behavior: 
We can see that for UMM there 
is a (close to) linear decrease of the fitness of new solutions over each new
evaluation. This suggests that UMM is able to model the fitness
landscape. Regarding CEGO, \ekhine{we observe a similar behavior to UMM for the first
50-100 evaluations, however, in several instances CEGO reaches the time limit
of 144 hours, so no observations can be made for longer budgets. In smaller
instances (e.g., Figure~\ref{fig:lolib}, top), CEGO again appears to stagnate.}{ the objective function decrease slowly?}

% \ekhine{In the case of the PFSP, the progress of both algorithms resembles a random search, as illustrated by the results
%  obtained on \texttt{rec05} (Fig.~\ref{fig:rec05}), which is the smallest PFSP
% instance considered here. The lack of any apparent convergence suggests that
% the models (both Bayesian and probabilistic) are not learning anything about
% the fitness landscape. Results for larger PFSP corroborate these conclusions (\supplement).}{}
% \manuel{Although previous studies~\citep{ZaeStoBar2014:ppsn} concluded
% that other distance measures are more suited than Kendall's-$\tau$ distance for
% the PFSP, the reported fitness differences between various distance metrics are
% small.
% Hence, we conjecture that the same behavior will be observed with other distance metrics.}{}\MANUEL{On second thought, I'm not sure we should say this, because we really have no idea and we actually hope it is not true.}

@@

>>>>

\usepackage{booktabs} % For formal tables
\usepackage{enumitem}
\usepackage{afterpage,tabularx}
\hypersetup{draft}
%\usepackage[colorlinks=true, allcolors=blue]{hyperref}

@@

\newcommand{\CEGOorig}{CEGO$_\text{ori}$\xspace}
\newcommand{\CEGOinv}{CEGO$_\text{inv}$\xspace}

\subsection{Rankings and orderings: experiments}\label{sec:exper_ro}

In Section~\ref{sec:permutations}, we highlighted the important difference between rankings and orderings, usually neglected in the optimization literature. Statistical models (Borda or the \ken distance) are usually defined for rankings while the optimization literature usually considers orderings. An example of such confusion is the claim that Swap distance is the same metric as  the \ken distance, which is false except for particular permutations. In reality, the \emph{swap} distance is the \ken distance of the inverted permutations.

As a consequence of this important distinction, when evaluating classical
combinatorial problems, we must convert rankings to orderings by inverting the
permutation, as done in Eq.~\ref{eq:lop} and Eq.~\ref{eq:pfsp}, for the LOP and
PFSP, respectively. This might seem an implementation detail but a failure to use the correct interpretation has grave consequences.

During our experiments, we realized that the \emph{swap} distance used by CEGO
assumes rankings, i.e., it actually measures the \ken distance, yet CEGO did
not invert the permutation before evaluating the PFSP. We verified that this is
indeed how the implementation works and that those results match the published
ones~\citep{ZaeStoBar2014:ppsn}. Of course, the obvious question arises of what
would be the effect of inverting the permutation before calling the objective
function, so that the rankings are transformed to ordering, as done by our UMM
algorithm. Therefore, in the following, we show results of both variants, that is, the original CEGO (\CEGOorig) and our modified version that inverts the permutation before evaluation (\CEGOinv).

\manuel{}{The above replaces the text below?}

\ekhine{}{este primer parago es bastante importante para dejarlo aqui. }
In Section~\ref{sec:preliminaries} we highlighted the great difference of rankings and orderings, usually ommitted in the literature. In this lines, we show the consequences of misinterpreting permutations. We remind the reader that statistical models (Borda or the \ken distance) are defined for rankings while standard literature in optimization considers orderings. This is the reason while in the optimization field we will usually find the Swap distance instead of the \ken distance. However, both are usually claimed to be the same, which is false: they are the same in particular cases, for example when  one of the permutations is the identity or an inversion.

In Section~\ref{sec:permutations} we highlighted the great difference of rankings and orderings, usually ommitted in the literature. In this lines, we show the consequences of misinterpreting permutations. We remind the reader that statistical models (Borda or the \ken distance) are defined for rankings while standard literature in optimization considers orderings. This is the reason while in the optimization field we will usually find the Swap distance instead of the \ken distance. However, both are usually claimed to be the same, which is false: they are the same in particular cases, for example when  one of the permutations is the identity or an inversion.

@@

The red

@@

\paragraph{Discussion on other problems: the case of QAP}
